---
title: "HW03-121090230-蒋飞弘"
author: "Jiang Feihong_121090230"
date: "2024/11/05"
output:
  html_document: default
---
## a. Read Data
```{r}
yc=read.csv('D:/2025 Term 1/FIN4120 BY me/HW3/yc(2).csv')
print(head(yc))
yc=yc[,-1]
```

## b. Plot Data
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
yc1 = yc %>% filter(DATE == '2010-01-04'| DATE == '2019-06-28')
colnames(yc1) = c('DATE',1,3,6,12,24,36,60,84,120,240,360)
yc1.long = gather(yc1,rate,value, -DATE)
yc1.long$DATE = as.character(yc1.long$DATE)
yc1.long$rate = as.numeric(yc1.long$rate)
ggplot(yc1.long,aes(x=rate, y=value, group=DATE)) +
  geom_line(aes(color=DATE))+
  geom_point(aes(color=DATE)) +
  labs(title = 'Yield Curves', x='Maturity(months)', y='Interest Rate(%)')
```

### The shape of the curve on 2010-01-04 is increasing and the shape of the curve on 2019-06-28 is humped.
### The humped one is likely to indicate that the economy might enter into recession as in the short-term the interest rate is decreasing, which shows the pessimistic expectations of the market.

## c. Explainations Can be from Different Theories:

### 1) As the term structure reflects at a given time the market expectation of future short-term rates, if people think there will be economy recession in short term, the interest rate will decline in short time and recover in long run.
### 2) As Investors request risk premium for holding riskier long term bond, with the maturity being longer investors bear more uncertainty, investors require higher compensation for investing in long term bond–risk premium, leading to a upward slope yield curve.
### 3) As different investor categories only invest in certain segments of yield curve in accordance with their liabilities, so it depends on the supply curves and demand curves of long-term and short-term bonds. During growth periods, banks sell bond securities in order to meet corporations’ and individuals’ strong demand for loans, hence the relative rise in short-term yields compared to long-term yields.

## d. Correlation Matrix
```{r}
colnames(yc) = c('DATE',1,3,6,12,24,36,60,84,120,240,360)
rownames(yc)=yc$DATE
yc=yc[,2:ncol(yc)]
yclag = yc[2:nrow(yc),]
d_yc = yclag-yc[1:nrow(yc)-1,]
corr_m = cor(d_yc)
corr_m
```

## e. Plot Correlation
```{r}
corr_df = as.data.frame(corr_m)
corr_df = corr_df %>%
  filter(corr_df$`1`==1|corr_df$`360`==1) %>%
  mutate(name = c(1,360))
corr_df.long = gather(corr_df,maturity,value,-name)
corr_df.long$name = as.character(corr_df.long$name)
corr_df.long$maturity = as.numeric(corr_df.long$maturity)
ggplot(corr_df.long,aes(x=maturity, y=value, group=name)) +
  geom_line(aes(color=name))+
  geom_point(aes(color=name)) +
  labs(title = 'Correlation curve', x='Maturity(months)', y='Correlation')

```

### They are consistent as 1-month rate daily changes have higher correlation between short-term bonds and 30-year rate daily changes have higher correlation between long-run bonds.

## f. PCA
```{r}
corr = as.data.frame(corr_m)
corr = corr[4:11,4:11]
colnames(corr) = c('Y1','Y2','Y3','Y5','Y7','Y10','Y20','Y30')
rownames(corr) = c('Y1','Y2','Y3','Y5','Y7','Y10','Y20','Y30')


cov_m = cov(d_yc)
cov_m = cov_m[4:11,4:11]
ee=eigen(cov_m,symmetric=TRUE)
a = ee$values/sum(ee$values)
a[1]
a[1]+a[2]+a[3]


```

### We can find that the percentage of variations of yield curve changes can be explained by the first principal component is 87.47% and the percentage of variations of yield curve changes that can be explained by the first three principal components is 97.63%.

## g. Factor2
```{r}
plot(ee$vectors[,1],
     ylim=range(ee$vectors),
     type="l",xlab="tenor.id",ylab="",main="First 3 PCs of Interest Rates",lwd=2)
lines(ee$vectors[,2],
      col="darkblue",xlab="tenor.id",ylab="",lwd=2)
lines(ee$vectors[,3],
      col="lightblue",xlab="tenor.id",ylab="",lwd=2)
abline(h=0)
legend("topright",
       c("PC1","PC2","PC3"),lty=c(1,1,1),
       lwd=c(2,2,2),
       col=c("black","darkblue","lightblue"))
```

### As shown above, the second component is associated with slope oscillation or a measure ofthe steepness ofthe interest-rate curve. So if facotr2 > 0, in short run the yield is down and in long run the yield is up, which make the curve steeper.

## h. Factor3

### As shown in previous question, the third component has a different impact on each of the three segments of the yield curve.  It brings more or less concavity to the intermediate segment of the curve. So if facotr3 > 0, in short run the yield is up, in medium run the yield is down and in long run the yield is up, which make the concave and convex butterfly movements to this curve and make it medium part concave below and both ends upwards.

## i. Three Factors Model
```{r}
d_yc.1 = d_yc[,4:11]
ft = as.matrix(d_yc.1) %*% ee$vectors
tmp = lm(d_yc.1$`12`~0+ft[,1]+ft[,2]+ft[,3])
tmp$coefficients
tmp2 = lm(d_yc.1$`360`~0+ft[,1]+ft[,2]+ft[,3])
tmp2$coefficients
  
  
```

### The change in the 1 year maturity rate is -0.1988% and the change in the 30 year maturity rate is 0.4861%.

## j.Sensitivity
```{r}
sens=t(apply(d_yc.1,2,function(x) lm(x~0+ft[,1]+ft[,2]+ft[,3])$coefficients))

plot(sens[,1],type="l",ylim=range(sens),lwd=2,xlab="maturity",ylab="sensitivity",main="sensitivity")
lines(sens[,2],col="darkblue",lwd=2,xlab="maturity",ylab="sensitivity")
lines(sens[,3],col="blue",lwd=2,xlab="maturity",ylab="sensitivity")
abline(h=0)
legend("top",
       c("PC1","PC2","PC3"),lty=c(1,1,1),
       lwd=c(2,2,2),
       col=c("black","darkblue","blue"))

```

#### 1) As the data indicate that the component corresponding to the largest eigenvalue is such that 
$c_{11}$ ≈ $c_{12}$ ≈ $c_{13}$... ≈ $c_{1m}$

#### so it is associated with parallel movements in the interest-rate curve.

#### 2) As the component corresponding to the second largest eigen-value is such that either
$c_{21}$ ≥ $c_{22}$ ≥ $c_{23}$... ≥ $c_{2m}$

#### or
$c_{21}$ ≤ $c_{22}$ ≤ $c_{23}$... ≤ $c_{2m}$

#### This function crosses the x-axis for an interest rate corresponding to a maturity,  depending on the period and the country under consideration. This is the reason this component is associated with slope oscillation or a measure ofthe steepness ofthe interest-rate curve.

#### 3) The componentcorresponding to the third largest eigenvalue has a different impact on each of the three segments of the yield curve (short, medium and long term). It brings more or less concavity to the intermediate segment of the curve. Hence, for intermediate values of k, c3k is significantly lower or greater than the remaining values of k. This is the reason this component is associated with the curvature of the interest-rate curve.