---
title: "HW02_FIN4120"
author: "Jiang Feihong_121090230"
date: "2024/10/21"
output: html_document
---
# Q1

### a. 
$(1+\dfrac{R_{2}(0,1)}{2})^2$ = (1+$\dfrac{R_{2}(0,0.5)}{2}$)(1+$\dfrac{F_{2}(0,0.5,0.5)}{2}$)\
$1.02625^2$ = 1.025(1+$\dfrac{F_{2}(0,0.5,0.5)}{2}$)\
$F_{2}(0,0.5,0.5)$ = 5.5003%

```{r}

```


### b.
$(1+\dfrac{R_{2}(0,1.5)}{2})^3$ = (1+$\dfrac{R_{2}(0,0.5)}{2}$)$(1+\dfrac{F_{2}(0,0.5,1)}{2})^2$\
$1.02875^3$ = 1.025$(1+\dfrac{F_{2}(0,0.5,1)}{2})^2$\
$F_{2}(0,0.5,1)$ = 6.1260%

### c. The cash flows are coupons of 5% in six months(5) and a year, and coupon plus principal payment of 105% (105) in 18 months. We can discount using the spot rates that we are given:
P = $\dfrac{5}{(1+\dfrac{0.05}{2})}$+$\dfrac{5}{(1+\dfrac{0.0525}{2})^2}$+$\dfrac{105}{(1+\dfrac{0.0575}{2})^3}$ = 106.0661


# Q2

$R(0,1)$ = 0.07(first year spot rate)\
$R(0,3)$ = 0.09(two year spot rate)\
$F(0,1,2)$ = $F(0,1,3-1)$ = $[\dfrac{(1+R(0,3))^3}{(1+R(0,1))^1}]^\dfrac{1}{3-1}$-1 = 0.1001\
P = $\dfrac{1}{(1+F(0,1,3))^2}$ = 0.8263

# Q3

### a. The par yield c(n) for maturity n is given by the formula
$c(n)$ = $\dfrac{1-\dfrac{1}{(1+R(0,n))^n}}{\sum_{i=1}^{n}\dfrac{1}{(1+R(0,i))^i}}$

### From this equation, we obtain the following par yields:
```{r}
fuc = function(n,r){
  r = c(0.04,0.055,0.059,0.062,0.06382,0.0655,0.0665,0.06741,0.0683,0.069)
  a = 0
  for (i in (1:n)) {
    a = a + 1/(1+r[i])^i
  }
  c = (1-1/(1+r[n])^n)/a
  print(c)
}
fuc(1)
fuc(2)
fuc(3)
fuc(4)
fuc(5)
fuc(6)
fuc(7)
fuc(8)
fuc(9)
fuc(10)
```

### b. F(0,x,y − x), the forward rate as seen from date t = 0, starting at date t = x, and with residual maturity y − x is defined as
$F(0,x,y-x)$ =  $[\dfrac{(1+R(0,y))^y}{(1+R(0,x))^x}]^\dfrac{1}{y-x}$-1

### From the previous equation, we obtain the forward yield curve in one year
```{r}
fuc.1 = function(n,r){
  r = c(0.04,0.055,0.059,0.062,0.06382,0.0655,0.0665,0.06741,0.0683,0.069)
  f = ((1+r[n+1])^(n+1)/(1+r[1]))^(1/n)-1
  print(f)
}
fuc.1(1)
fuc.1(2)
fuc.1(3)
fuc.1(4)
fuc.1(5)
fuc.1(6)
fuc.1(7)
fuc.1(8)
fuc.1(9)
```



# Q4

### a. Recall that the par yield c(n) for maturity n is given by the formula
$c(n)$ = $\dfrac{1-\dfrac{1}{(1+R(0,n))^n}}{\sum_{i=1}^{n}\dfrac{1}{(1+R(0,i))^i}}$

### Using this equation, we obtain the following par yields:
```{r}
fuc = function(n,r){
  r = c(0.07,0.068,0.0662,0.0646,0.0633,0.0625,0.062,0.0616,0.06125,0.061)
  a = 0
  for (i in (1:n)) {
    a = a + 1/(1+r[i])^i
  }
  c = (1-1/(1+r[n])^n)/a
  print(c)
}
fuc(1)
fuc(2)
fuc(3)
fuc(4)
fuc(5)
fuc(6)
fuc(7)
fuc(8)
fuc(9)
fuc(10)
```

### b. Recall that F(0,x,y − x), the forward rate as seen from date t = 0, starting at date t = x, and with residual maturity y − x is defined as
$F(0,x,y-x)$ =  $[\dfrac{(1+R(0,y))^y}{(1+R(0,x))^x}]^\dfrac{1}{y-x}$-1

### Using the previous equation, we obtain the forward yield curve in one year
```{r}
fuc.1 = function(n,r){
  r = c(0.07,0.068,0.0662,0.0646,0.0633,0.0625,0.062,0.0616,0.06125,0.061)
  f = ((1+r[n+1])^(n+1)/(1+r[1]))^(1/n)-1
  print(f)
}
fuc.1(1)
fuc.1(2)
fuc.1(3)
fuc.1(4)
fuc.1(5)
fuc.1(6)
fuc.1(7)
fuc.1(8)
fuc.1(9)
```


# Q5
```{r}
data = c()
theta = 1
mu = 1.2
sigma = 0.3
z = rnorm(501,0,0.1)
x = rep(0,501)
for (n in -20:20) {
  x[1] = n
  for (i in 2:501){
    x[i] = x[i-1] + theta*(mu-x[i-1])*0.01 + sigma*z[i]
  }
  data = cbind(data,x)
}
data  = as.data.frame(data)
for (m in 1:ncol(data)) {
  names(data)[m] = m
}
library(tidyr)
library(dplyr)
data.long = gather(data,group,value)
data.long = data.long %>% group_by(group) %>%
  mutate(time = c(1:501)) %>%
  ungroup()
data.long$group = as.factor(data.long$group)
library(ggplot2)
ggplot(data.long, aes(x = time, y = value, color = group))+
  geom_line(size = 1)+geom_hline(yintercept= mu)


```


# Q6
```{r}
data = c()
theta = 1
mu = 1.2
sigma = 0.3
z = rnorm(1001,0,0.1)
x = rep(0,1001)
for (n in seq(from = 0, to = 4, by = 0.1)) {
  x[1] = n
  for (i in 2:1001){
    if(x[i-1]>0){
      x[i] = x[i-1] + theta*(mu-x[i-1])*0.01 + sigma*z[i]*x[i-1]^0.5
    } else{
      x[i] = x[i-1] + theta*(mu-x[i-1])*0.01
    }
  }
  data = cbind(data,x)
}
data  = as.data.frame(data)
for (m in 1:ncol(data)) {
  names(data)[m] = m
}
library(tidyr)
library(dplyr)
data.long = gather(data,group,value)
data.long = data.long %>% group_by(group) %>%
  mutate(time = c(1:1001)) %>%
  ungroup()
data.long$group = as.factor(data.long$group)
library(ggplot2)
ggplot(data.long, aes(x = time, y = value, color = group))+
  geom_line(size = 1)+geom_hline(yintercept= mu)

```

