---
title: "HW05"
author: "Feihong Jiang_121090230"
date: "2024/11/20"
output: html_document
---
# Q1

```{r}

yc = read.csv('D:/2025 Term 1/FIN4120 BY me/HW5/price_ttm.csv')
yc$price_09.26.2017 = as.numeric(yc$price_09.26.2017)/100
yc$yeild = -log(yc$price_09.26.2017)/yc$ttm
tau = 3.33
theta = yc$ttm

sens0 = 1
sens1 = (1-exp(-theta/tau))/(theta/tau)
sens2 = (1-exp(-theta/tau))/(theta/tau)-exp(-theta/tau)
tmp = data.frame(yc=yc$yeild,
               sens0=sens0,
               sens1=sens1,
               sens2=sens2)

rst = lm(tmp$yc ~ tmp$sens1+tmp$sens2)
tmpcoef = rst$coefficients

tenors = seq(0.01,30,0.01)
term0 = tmpcoef[1]
term1 = tmpcoef[2]*(1-exp(-tenors/tau))/(tenors/tau)
term2 = tmpcoef[3]*((1-exp(-tenors/tau))/(tenors/tau)-exp(-tenors/tau))
fitted_yield_curve = term0+term1+term2
fit_yc = data.frame(yc=fitted_yield_curve,
                  tenor=tenors)

plot(fitted_yield_curve~tenors,type='l',
     ylim = c(0,0.03),xlim = c(0,30),xlab = "tenor", ylab = "rate", main = "Fitted Yield Curve")
par(new=T)
plot(yc$yeild~theta,type='l',col="red",
     ylim = c(0,0.03),xlim = c(0,30),xlab = "",ylab = "")
print(fit_yc$yc[which(fit_yc$tenor==5)])
```

# Q3

## a. Function
```{r}
Duration = function(pr,cpr,ttm,ytm){
  cp = pr*cpr
  q = 1/(1+ytm)
  cfc = sum(cp*q^seq(1,ttm))
  cfp = pr*q^ttm
  price = cfc + cfp
  
  avgtime = sum(cp*q^seq(1,ttm)*seq(1,ttm)) + pr*q^ttm*ttm
  macdur = avgtime/price
  moddur = macdur*q
  doldur = moddur*price
  DV01 = doldur/10000
  
  rst=list(price=price, macaulay_duration=macdur, modified_duration=moddur, dollar_duration=doldur, DV01=DV01)
  return(rst)
}
```

## b. Calculation
```{r}
B1=c()
y_c = c(0.05,0.07,0.0799,0.0801,0.09,0.11)
for (i in 1:6) {
  x = c(YTM=y_c[i],
               prc=as.numeric(Duration(100,0.08,25,y_c[i])[1]),
               McD=as.numeric(Duration(100,0.08,25,y_c[i])[2]),
               MdD=as.numeric(Duration(100,0.08,25,y_c[i])[3]),
               DD=as.numeric(Duration(100,0.08,25,y_c[i])[4]),
               DV01=as.numeric(Duration(100,0.08,25,y_c[i])[5]))
  B1 = rbind(B1,x)
}
B1 = as.data.frame(B1)
B1$name = "bond1"

B2=c()
for (i in 1:6) {
  x = c(YTM=y_c[i],
               prc=as.numeric(Duration(100,0.08,5,y_c[i])[1]),
               McD=as.numeric(Duration(100,0.08,5,y_c[i])[2]),
               MdD=as.numeric(Duration(100,0.08,5,y_c[i])[3]),
               DD=as.numeric(Duration(100,0.08,5,y_c[i])[4]),
               DV01=as.numeric(Duration(100,0.08,5,y_c[i])[5]))
  B2 = rbind(B2,x)
}
B2 = as.data.frame(B2)
B2$name = "bond2"

B3=c()
for (i in 1:6) {
  x = c(YTM=y_c[i],
               prc=as.numeric(Duration(100,0.05,25,y_c[i])[1]),
               McD=as.numeric(Duration(100,0.05,25,y_c[i])[2]),
               MdD=as.numeric(Duration(100,0.05,25,y_c[i])[3]),
               DD=as.numeric(Duration(100,0.05,25,y_c[i])[4]),
               DV01=as.numeric(Duration(100,0.05,25,y_c[i])[5]))
  B3 = rbind(B3,x)
}
B3 = as.data.frame(B3)
B3$name = "bond3"

B4=c()
for (i in 1:6) {
  x = c(YTM=y_c[i],
               prc=as.numeric(Duration(100,0.05,5,y_c[i])[1]),
               McD=as.numeric(Duration(100,0.05,5,y_c[i])[2]),
               MdD=as.numeric(Duration(100,0.05,5,y_c[i])[3]),
               DD=as.numeric(Duration(100,0.05,5,y_c[i])[4]),
               DV01=as.numeric(Duration(100,0.05,5,y_c[i])[5]))
  B4 = rbind(B4,x)
}
B4 = as.data.frame(B4)
B4$name = "bond4"

Bond = rbind(B1,B2,B3,B4)

library(tidyr)
library(dplyr)
prc = Bond %>%
  select(YTM,name,prc)
prc = spread(prc,name,prc)

McD = Bond %>%
  select(YTM,name,McD)
McD = spread(McD,name,McD)

MdD = Bond %>%
  select(YTM,name,MdD)
MdD = spread(MdD,name,MdD)

DD = Bond %>%
  select(YTM,name,DD)
DD = spread(DD,name,DD)

DV01 = Bond %>%
  select(YTM,name,DV01)
DV01 = spread(DV01,name,DV01)

prc
McD
MdD
DD
DV01
```

## c. Conclusion

#### YTM: Macaulay Duaration, Modified Duration and Dollar Duration will all decrease when YTM is increasing.
#### Coupon Rate: Macaulay Duaration and Modified Duration will both decrease when coupon rate is increasing, while Dollar Duration will increase when coupon rate is increasing.
#### TTM: Macaulay Duaration, Modified Duration and Dollar Duration will all decrease when YTM is decreasing.

## d. Amend function
```{r}
Duration_1 = function(pr,cpr,ttm,ytm,freq){
  cp = pr*cpr/freq
  q = 1/(1+ytm/freq)
  cfc = sum(cp*q^seq(1,ttm*freq))
  cfp = pr*q^(ttm*freq)
  price = cfc + cfp
  
  avgtime = sum(cp*q^seq(1,ttm*freq)*seq(1/freq,ttm,by=1/freq)) + pr*ttm*q^(ttm*freq)
  macdur = avgtime/price
  moddur = macdur*q
  doldur = moddur*price
  DV01 = doldur/10000
  
  rst=list(price=price, macaulay_duration=macdur, modified_duration=moddur, dollar_duration=doldur, DV01=DV01)
  return(rst)
}

Duration_1(100,0.2,2,0.04,2)
```

## e. Continuous
```{r}
Duration_con = function(pr,cpr,ttm,ytm){
  cp = pr*cpr
  cfc = sum(cp*exp(-seq(1,ttm)*ytm))
  cfp = pr*exp(-ttm*ytm)
  price = cfc + cfp
  
  avgtime = sum(cp*exp(-seq(1,ttm)*ytm)*seq(1,ttm)) + cfp*ttm
  macdur = avgtime/price
  moddur = macdur
  doldur = moddur*price
  DV01 = doldur/10000
  
  rst=list(price=price, macaulay_duration=macdur, modified_duration=moddur, dollar_duration=doldur, DV01=DV01)
  return(rst)
}
Duration_con(100,0.1,2,0.04)[2]
Duration_con(100,0.1,2,0.04)[3]
```

